Ktirio Urban Building: A Computational Framework for City Energy Simulations Enhanced by CI/CD Innovations on EuroHPC Systems

### 2024-10-26 article.hid2-ub-cicd.ppam24.tex

We now turn to describing the physical modeling of city energy simulation.
To do this, we built on two main tools: Feel++ and Modelica.

@@

>>>>

We now turn to describing the physical modeling of city energy simulation.

To do this, we built on two main tools: Feel++ and Modelica. 
\subsubsection{Governing equation of the simulation}
The physical model of the city is based on a multizone model. One zone is allocated per building floor. The energy balance for each zone is calculated taking into account energy exchanges with adjacent zones and the environment.
The energy balance is written according to the first law of thermodynamics : in an isolated system, the sum of all forms of energy is constant.
\begin{equation}
    C\cdot \frac{dT}{dt}=\phi_{HVAC}+\phi_{gain}-\phi_{losses}
\end{equation}
To give an overview of the equation's term we will give a brief summary.
\begin{itemize}
  \item $C\cdot \frac{dT}{dt}$ : represents the thermal inertia of the construction. 
  This term quantifies how the temperature of the zone changes over time due to its thermal mass. 
  \begin{itemize}
    \item $C$ :This is a measure of how much heat a material can store and release. 
    It combines the heat capacity of the construction (how much energy is required to change the temperature) with the rate of temperature change.
    \item $\frac{dT}{dt}$ : This is the rate of change of temperature with respect to time. 
    A higher thermal inertia means the construction will have a slower rate of temperature change in response to energy inputs and losses, effectively buffering temperature fluctuations.
  \end{itemize}
  \item $\phi_{HVAC}$ : This term represents the energy input from the Heating, Ventilation, and Air Conditioning (HVAC) systems. 
  It includes energy added to the zone to maintain desired temperature levels.
  \item $\phi_{gain}$ : Referred to as free energy gain, this term includes solar gains (heat from sunlight) and internal gains (heat generated by occupants and equipment).
  \item $\phi_{losses}$ : Represents the energy lost through walls, roofs, floors, and air exchanges with the environment.
\end{itemize}

\textbf{HVAC heatflow}

In this model, the HVAC system is treated as an ideal heat source, which means it adjusts its output to maintain a set temperature within the building.

\textbf{Heat gain}

Heat gain is the addition of energy to the building, derived from two sources: \textbf{internal heat gains} and \textbf{solar heat gains}.\\

\textbf{Internal heat gain}

The internal heat gain captures the heat produced by occupants inside the building. 
It is modeled as a function of metabolic rate $Met$ and the number of inhabitants, which varies according to a timetable.
\begin{equation}
  \phi_{int}=Met\cdot d_{inhabitant}(timetable)
\end{equation}

\textbf{Solar heat gain}

Calculate the total solar heat gain, which includes contributions from direct, diffuse, and reflected solar radiation, taking into account the building’s surface characteristics and shading coefficient..
\begin{equation}
  \phi_{solar}=\alpha\cdot S\cdot(E_{dir}+E_{diff}+E_{ref})\cdot S_C
\end{equation}
Direct solar radiation is the component of solar radiation that is adjusted for the angle of incidence of the sunlight.
\begin{equation}
  E_{dir}=F_{dir}\cdot cos(I)
\end{equation}
$F_{dir}$ is the direct solar radiation it is a function of the horizontal solar radiation measured by a weather station.
\begin{equation}
  F_{dir}=\frac{F_{dir,Hz}}{sin(H_s)}
\end{equation}
$Cos (I)$ : calculates the cosine of the angle of incidence of sunlight, incorporating the sun's altitude and azimuth angles as well as the building's tilt angle
\begin{equation}
  cos(I)=cos(H_s)\cdot sin(\beta)\cdot cos(A_{zs}-A_{z})+sin(H_s)\cdot cos(\beta)
\end{equation}
Diffuse solar radiation, which includes contributions from the sky dome and is adjusted for the building's tilt.
\begin{equation}
  E_{diff}=F_{dif,Hz}+\frac{1+cos(\beta)}{2}
\end{equation}
Reflected solar radiation, calculated using the total horizontal radiation and the environnement's albedo
\begin{equation}
  E_{ref}=[F_{dir,Hz}+F_{dif,Hz}]\cdot \frac{1+cos(\beta)}{2}\cdot albedo
\end{equation}

\textbf{Conduction losses}

Energy loss through walls is computed by summing the losses across all wall segments.
It depends on the heat transfer coefficient, the wall area, and the temperature difference between the indoor and outdoor environments.

In our current model, we simplify the calculations by not including the effects of thermal bridges, as they depends of specifics informations on the building's envelope, which we do not have.
\begin{equation}
  \phi_{wall}=\sum_{j} U_j\cdot S_j\cdot \Delta T
\end{equation}

The convection losses are computed differently regarding the surface of the wall they are applied to.

On the interior surface we apply a constant :
\begin{equation}
  \phi_{conv,int}=\alpha_{conv} \cdot \Delta T
\end{equation}

On the exterior surface we use a forced convection model :
\begin{equation}
  \phi_{conv,ext}=(1.8+4.8\cdot v_{air}) \cdot \Delta T
\end{equation}

\textbf{Radiation losses}

Radiation losses describe the heat energy lost due to thermal radiation. 
This is an important factor in buildings as it impacts both energy efficiency and thermal comfort.
We differentiate between the exterior of the wall that exchange with the sky and the interior where the walls exchange between each other.

Exterior radiation exchange. 
\begin{equation}
  \phi_{net,sky} = \varepsilon_i \cdot \sigma \cdot (T_i^4-T_{sky}^4)
\end{equation}
Interior radiation exchange.
\begin{equation}
  J[i] = \varepsilon_i \cdot \sigma \cdot T_i^4 + (1.0 - \varepsilon_i) \cdot \sum_{j=1}^{n} F[i,j] \cdot J[j]
\end{equation}
Radiative heat exchange $J[i]$ :
\begin{itemize}
  \item $\varepsilon_i \cdot \sigma \cdot T_i^4$ : 
  This term calculates the heat radiated by the surface i. 
  $\epsilon$ is a measure of how effectively the surface emit radiation.
  The $T^4$ reflect the fact that radiative heat loss increases rapidly with temperature.
  \item $(1.0 - \varepsilon_i) \cdot \sum_{j=1}^{n} F[i,j] \cdot J[j]$ : 
  This term accounts for the heat absorbed from other surfaces $j$ within the zone.
  The factor $(1.0 - \varepsilon_i)$ adjusts for the fraction of incoming radiation that is not emitted by surface $i$.
\end{itemize}
\begin{equation}
  \phi_{net,int}[i] = S\cdot \frac{\varepsilon_i}{1.0 - \varepsilon_i} \cdot (\sigma \cdot T_i^4 - J[i])
\end{equation}
Net radiation flow $Q_{flow}[i]$ :
\begin{itemize}
  \item $\frac{\varepsilon_i}{1.0 - \varepsilon_i} \cdot (\sigma \cdot T_i^4 - J[i])$ :
  This term calculate the net heat flow due to radiation for surface $i$.
  It considers the difference between the heat emited by the surface and the heat received from other surfaces.
  The ratio $\frac{\varepsilon_i}{1.0 - \varepsilon_i}$ scales the net flow based on the emissivity of the surface.
\end{itemize}

\textbf{Air change}

This term represents the energy loss due to air exchange between the zone and the exterior.
\begin{equation}
  \phi_{air}=0.34\cdot Q_v\cdot \Delta T
\end{equation}

@@

\textbf{Nomenclature}\\
$\alpha$ : Absorption of a material.\hfill[-]\\
$\alpha_{conv}$=10 : heat conductance of convection\hfill[-]\\
$\beta$ : Angle of inclination of the wall to the horizontal.\hfill[-]\\
$\Delta T$ : Temperature difference.\hfill[K]\\
$\epsilon$ : Emissivity of the surface.\hfill[-]\\
$\phi_{air}$ : Air heatflow.\hfill[W]\\
$\phi_{net}$ : radiative heatflow.\hfill[W]\\
$\phi_{solar}$ : Solar heatflow from short wave lenght.\hfill[W]\\
$\phi_{wall}$ : Heatflow through a wall.\hfill[W]\\
$\sigma$ : Stefan-Boltzmann constant $5,67\cdot10^{-8}$\hfill[W$\cdot m^{-2}\cdot K^{-4}$]\\
$albedo$ : Reflective properties of the material.\hfill[-]\\
$A_z$ : Azimuth of the surface.\hfill[-]\\
$A_{zs}$ : Azimuth of the sun.\hfill[-]\\
$C$ : Heatcapacity of the zone.\hfill[W$\cdot K^{-1}$]\\
$d_{inhabitant}(timetable)$ : density of inhabitant in a building as a function of a timetable set for the building type.\hfill[-]\\
$E_{dir}$ : Direct solar heatflow.\hfill[W$\cdot m^{-2}$]\\
$E_{diff}$ : Diffuse solar heatflow.\hfill[W$\cdot m^{-2}$]\\
$E_{ref}$ : Reflected solar heatflow.\hfill[W$\cdot m^{-2}$]\\
$F_{dir}$ : Direct solar radiation.\hfill[W$\cdot m^{-2}$]\\
$F_{dir,Hz}$ : Direct solar radiation on a horizontal surface given by climate data.\hfill[W$\cdot m^{-2}$]\\
$F_{ij}$ : View factor between a $surface_{(i)}$ and a $surface_{(j)}$.\hfill[-]\\
$H_s$ : Angular height of the sun. \hfill[-]\\
$J$ : Radiosity of a surface.\hfill[$W\cdot m^{-2}$]\\
$Met$ : Metabolic rate\hfill[W]\\
$0.34\cdot Q_v$ : Airflow as a fraction of the zone's volume per hour multiplied by the air volumetric heat capacity

\raggedleft{[$W\cdot K^{-1}$]}\\
$S$ : Surface.\hfill[$m^2$]\\
$S_C$ : Shading coefficient $\in$[0,1]\hfill[-]\\
$U_j$ : Conductance of a $wall_j$.\hfill[$W\cdot m^{-2}\cdot K^{-1}$]\\
$V_{air}$ : Wind speed \hfill[m$\cdot s^{-1}$]
